---
title: "R Notebook"
output: html_notebook
---

Read in data
```{r}
Qref <- read.csv("GNOMADisec.20.Q", header = FALSE, sep = " ", stringsAsFactors = FALSE)
Qsam <- read.csv("ORIENisec_missing10.20.Q", header = FALSE, sep = " ", stringsAsFactors = FALSE)
Qboth <- rbind(Qref,Qsam)

fam_ref <- read.csv("GNOMADisec.fam", header = FALSE, sep = " ", stringsAsFactors = FALSE)
fam_sam <- read.csv("ORIENisec_missing10.fam", header = FALSE, sep = " ", stringsAsFactors = FALSE)
fam_both <- rbind(fam_ref,fam_sam)

names(fam_both)[1] <- "WES_ID"
Qfam2 <- cbind(Qboth,fam_both)
Qfam2 <- Qfam2[,c(21,1:20)]

df4 <- merge(Qfam2, IDs2, by="WES_ID")
data_tsne4 <- df4[,c(1,22,23,2:21)]

df <-subset(data_tsne4, data_tsne4$superpop != "N/A")
names(df)[4] <- "Q1"
names(df)[5] <- "Q2"
names(df)[6] <- "Q3"
names(df)[7] <- "Q4"
names(df)[8] <- "Q5"
names(df)[9] <- "Q6"
names(df)[10] <- "Q7"
names(df)[11] <- "Q8"
names(df)[12] <- "Q9"
names(df)[13] <- "Q10"
names(df)[14] <- "Q11"
names(df)[15] <- "Q12"
names(df)[16] <- "Q13"
names(df)[17] <- "Q14"
names(df)[18] <- "Q15"
names(df)[19] <- "Q16"
names(df)[20] <- "Q17"
names(df)[21] <- "Q18"
names(df)[22] <- "Q19"
names(df)[23] <- "Q20"
write.csv(df, "ORIEN_gnomAD_K20_Admixture_Qvalues_with_IDs.csv")

```

Execute knn on a dataset and calculate the accuracy metrics for the predictions.
  If the input matrix number of columns does not match the number of values in `code`, then the matrix is transposed and used.

  m = The matrix of samples (columns) for a series of attributes (rows).
  codes = classes/codes to train to predict.
  only_accuracy = Return only accuracy (single value) or a list of metrics (default=TRUE).
  k = The k to use in knn
  return = Either a single number (accuracy) or a list (accuracy, extended metrics) if the parameter `only_accuracy` is set to FALSE.

Example:
simple_metric(t(iris[,1:4]), iris[,5])

-------------------------
RUN FOR SUPERPOPULATIONS
```{r}
simple_metric <- function(m, codes, only_accuracy=TRUE, k = 5) {
  if ( nrow(m) != length(codes))
    m <- t(m)
  stopifnot(nrow(m) == length(codes))

  class_labels <- unique(sort(codes))

  preds <- class::knn(m, m, cl = codes, k = k)

  cm <- caret::confusionMatrix(
    data=factor(preds, levels = class_labels),
    reference = factor(codes, levels = class_labels)
  )
  if ( only_accuracy )
    cm$overall[["Accuracy"]]
  else
    list(accuracy = cm$overall[["Accuracy"]], summary = cm)
}

simple_metric(t(df[,4:23]), df[,3])
```

-------------------------
RUN FOR SUBPOPULATIONS
```{r}
simple_metric(t(df[,4:23]), df[,2])
```

