---
title: "Admixture_Stacked_Barplot"
output: html_notebook
---

Load libraries
```{r}
library(ggplot2)
library(reshape2)
library(gridExtra)
library(dplyr)
library(data.table)
library(tidyverse)
library(plotly)
library(plyr)
library(stringi)
```

Read in reference population info from IGSR, race/ethnicity info from ORIEN and TCGA.
```{r}

clinlinkage <- read.csv("20PRJ071USC_20210928_ClinicalMolLinkage_V4.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
key_SLID <- subset(clinlinkage, Tumor.Germline == "Germline")
key_SLID2 <- key_SLID[,c(1,3)]
demos <- read.csv("20PRJ071USC_20210928_PatientMaster_V4.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
demos2 <- demos[,c(1,5,6)]
names(key_SLID2)[names(key_SLID2 ) == "ORIENAvatarKey" ] <- "AvatarKey"
merge <- merge(x = key_SLID2,y = demos2, by='AvatarKey', sort=FALSE, all=TRUE)
names(merge)[names(merge ) == "WES" ] <- "WES_ID"
merge2 <- merge[,c(2,3,4)]

USC <-  read.csv("20211203_USC_ClinicalMolLinkage_V4.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
names(USC)[names(USC ) == "ORIENAvatarKey" ] <- "key"
key <- read.csv("V4_RecordKey_Jan2022.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
names(key)[names(key ) == "Avatar.Patient.ID" ] <- "key"
key <- key[,c(4,6,7)]
USC2 <- merge(x=USC, y=key, by="key")
USC2 <- USC2[,c(4,21,22)]
names(USC2)[names(USC2 ) == "WES" ] <- "WES_ID"
all <- bind_rows(merge2,USC2)
all <- all[-which(all$WES_ID == ""), ]

all[all$Race == "Black",]$Race = "Black_or_African_American"
all[all$Race == "Unknown/Not Reported",]$Race = "Unknown/Not_Reported"
all[all$Race == "Other Asian, including Asian, NOS and Oriental, NOS",]$Race = "Asian"
all[all$Race == "Filipino",]$Race = "Asian"
all[all$Race == "Chinese",]$Race = "Asian"
all[all$Race == "Korean",]$Race = "Asian"
all[all$Race == "American Indian, Aleutian, or Eskimo (includes all indigenous populations of the Western Hemisphere)",]$Race = "American_Indian_or_Alaska_Native"
all[all$Race == "American Indian, Aleutian, or Eskimo (includes all indigenous populations of the Western hemisphere)",]$Race = "American_Indian_or_Alaska_Native"
all[all$Race == "Asian Indian or Pakistani, NOS",]$Race = "Asian"
all[all$Race == "Asian Indian",]$Race = "Asian"
all[all$Race == "Vietnamese",]$Race = "Asian"
all[all$Race == "Japanese",]$Race = "Asian"
all[all$Race == "Pacific Islander, NOS",]$Race = "Hawaiian_or_Other_Pacific_Islander"
all[all$Race == "Unknown/Not Applicable",]$Race = "Unknown/Not_Reported"
all[all$Race == "Hawaiian",]$Race = "Hawaiian_or_Other_Pacific_Islander"
all[all$Race == "Laotian",]$Race = "Asian"
all[all$Race == "Pakistani",]$Race = "Asian"
all[all$Race == "Kampuchean (Cambodian)",]$Race = "Asian"
all[all$Race == "Polynesian, NOS",]$Race = "Hawaiian_or_Other_Pacific_Islander"
all[all$Race == "Thai",]$Race = "Asian"
all[all$Race == "Guamanian, NOS",]$Race = "Hawaiian_or_Other_Pacific_Islander"

all[all$Ethnicity == "Non-Spanish; Non-Hispanic",]$Ethnicity = "Not_Hispanic/Not_Latino" 
all[all$Ethnicity == "Spanish, NOS; Hispanic, NOS; Latino, NOS",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Unknown/Not Reported",]$Ethnicity = "Unknown/Not_Reported"
all[all$Ethnicity == "Mexican (includes Chicano)",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "South or Central American (except Brazil)",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Puerto Rican",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Unknown/Not Applicable",]$Ethnicity = "Unknown/Not_Reported"
all[all$Ethnicity == "Cuban",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Other specified Spanish/Hispanic origin (includes European; excludes Dominican Republic)",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Spanish surname only",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Dominican Republic",]$Ethnicity = "Hispanic/Latino"


TCGA_race <- read.csv("TCGA_vcfIDs_race_uniq_final.tsv", sep = "\t", header = FALSE, stringsAsFactors = FALSE, fill = TRUE)
TCGA_ethnicity <- read.csv("TCGA_vcfIDs_ethnicity_uniq_final.tsv", sep = "\t", header = FALSE, stringsAsFactors = FALSE, fill = TRUE)
names(TCGA_race)[2] <- "Race"
names(TCGA_ethnicity)[2] <- "Ethnicity"
names(TCGA_race)[1] <- "WES_ID"
names(TCGA_ethnicity)[1] <- "WES_ID"
TCGA <- merge(x = TCGA_race, y = TCGA_ethnicity, by="WES_ID")

merge5 <- bind_rows(TCGA,all)

race <- merge5[,c(1,2)]
names(race)[names(race) == "Race"] <- "pop"
ethnicity <- merge5[,c(1,3)]
names(ethnicity)[names(ethnicity) == "Ethnicity"] <- "pop"

str_replace(race$WES_ID, "_st_g", "")
str_replace(ethnicity$WES_ID, "_st_g", "")

# https://www.internationalgenome.org/data-portal/data-collection/30x-grch38
# https://www.internationalgenome.org/data-portal/data-collection/hgdp

#refpops <- read.table("igsr_samples3.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
refpops_HGDP <- read.table("igsr-human genome diversity project.tsv.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
refpops_1000GP <- read.table("igsr-1000 genomes 30x on grch38.tsv.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
refpops_igsr <- read.table("igsr_samples3.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)

refpops_igsr <- refpops_igsr[,c(1,7)]
refpops_HGDP2 <- refpops_HGDP[,c(1,7)]
refpops_1000GP2 <- refpops_1000GP[,c(1,7)]
refpops2 <- unique(rbind(refpops_HGDP2,refpops_1000GP2,refpops_igsr))
names(refpops2)[names(refpops2) == "Sample.name"] <- "WES_ID"
names(refpops2)[names(refpops2) == "Superpopulation.name"] <- "pop"

```

Standardize/harmonize and combine fam + Q + pop info
```{r}
Q <- read.table("isec.5.Q", head = FALSE,  sep = " ", stringsAsFactors = FALSE)
raw_fam <- read.table("isec.fam", head = FALSE, sep = " ", stringsAsFactors = FALSE)
raw_fam$V1 <- sub("_st_g", "", raw_fam$V1)

raw_fam <- raw_fam[,c(1,2)]
names(raw_fam)[1] <- "WES_ID"
raw_fam$order <- 1:nrow(raw_fam) 
Q$order <- 1:nrow(Q) 

write.csv(race, "ORIEN_TCGA_IDs_race.csv")
write.csv(ethnicity, "ORIEN_TCGA_IDs_ethnicity.csv")
write.csv(refpops2, "1000GP_HGDP_IDs_Pops.csv")

HGDP_1000GP <- merge(x = raw_fam, y = refpops2, by="WES_ID")
HGDP_1000GP <- HGDP_1000GP[,c(3,1,4)]
#refs_numbered_ID_pop_Q <- cbind(raw_fam,Q)
refs_numbered_ID_pop_Q <- merge(HGDP_1000GP, Q, by="order" )
names(refs_numbered_ID_pop_Q)[4] <- "Q1"
names(refs_numbered_ID_pop_Q)[5] <- "Q2"
names(refs_numbered_ID_pop_Q)[6] <- "Q3"
names(refs_numbered_ID_pop_Q)[7] <- "Q4"
names(refs_numbered_ID_pop_Q)[8] <- "Q5"

Race <- merge(x = raw_fam, y = race, by="WES_ID")
samples_numbered_ID_race_Q <- cbind(raw_fam,Q)
samples_numbered_ID_race_Q <- merge(Race, Q, by="order" )
samples_numbered_ID_race_Q <- samples_numbered_ID_race_Q[,c(1,2,4:9)]
names(samples_numbered_ID_race_Q)[4] <- "Q1"
names(samples_numbered_ID_race_Q)[5] <- "Q2"
names(samples_numbered_ID_race_Q)[6] <- "Q3"
names(samples_numbered_ID_race_Q)[7] <- "Q4"
names(samples_numbered_ID_race_Q)[8] <- "Q5"

Ethnicity <- merge(x = raw_fam, y = ethnicity, by="WES_ID")
samples_numbered_ID_ethnicity_Q <- cbind(raw_fam,Q)
samples_numbered_ID_ethnicity_Q <- merge(Ethnicity, Q, by="order" )
samples_numbered_ID_ethnicity_Q <- samples_numbered_ID_ethnicity_Q[,c(1,2,4:9)]
names(samples_numbered_ID_ethnicity_Q)[4] <- "Q1"
names(samples_numbered_ID_ethnicity_Q)[5] <- "Q2"
names(samples_numbered_ID_ethnicity_Q)[6] <- "Q3"
names(samples_numbered_ID_ethnicity_Q)[7] <- "Q4"
names(samples_numbered_ID_ethnicity_Q)[8] <- "Q5"

write.csv(unique(refs_numbered_ID_pop_Q), "refs_numbered_ID_pop_Q.csv")
write.csv(unique(samples_numbered_ID_race_Q), "samples_numbered_ID_race_Q.csv")
write.csv(unique(samples_numbered_ID_ethnicity_Q), "samples_numbered_ID_ethnicity_Q.csv")

European <- subset(refs_numbered_ID_pop_Q, pop == "European Ancestry")
European_West_Eurasia_SGDP <- subset(refs_numbered_ID_pop_Q, pop == "European Ancestry,West Eurasia (SGDP)")
Europe_HGDP <- subset(refs_numbered_ID_pop_Q, pop == "Europe (HGDP)")
EUR <- rbind.fill(European, European_West_Eurasia_SGDP, Europe_HGDP)

East_Asian <- subset(refs_numbered_ID_pop_Q, pop == "East Asian Ancestry")
East_Asian_SGDP <- subset(refs_numbered_ID_pop_Q, pop == "East Asia (SGDP),East Asian Ancestry")
East_Asian_HGDP <- subset(refs_numbered_ID_pop_Q, pop == "East Asia (HGDP)")
East_Asian_SGDP2 <- subset(refs_numbered_ID_pop_Q, pop == "East Asia (SGDP)")
East_Asian_SGDP_HGDP <- subset(refs_numbered_ID_pop_Q, pop == "East Asia (SGDP),East Asia (HGDP)")
EAS <- rbind.fill(East_Asian, East_Asian_HGDP, East_Asian_SGDP, East_Asian_SGDP2, East_Asian_SGDP_HGDP)

American <- subset(refs_numbered_ID_pop_Q, pop == "American Ancestry")
America_HGDP <- subset(refs_numbered_ID_pop_Q, pop == "America (HGDP)")
America_SGDP <- subset(refs_numbered_ID_pop_Q, pop == "America (SGDP)")
America_SGDP_HGDP <- subset(refs_numbered_ID_pop_Q, pop == "America (SGDP),America (HGDP)")
AMR <- rbind.fill(America_HGDP, America_SGDP, America_SGDP_HGDP, American)

South_Asian <- subset(refs_numbered_ID_pop_Q, pop == "South Asian Ancestry")
South_Asia_SGDP <- subset(refs_numbered_ID_pop_Q, pop == "South Asia (SGDP),South Asian Ancestry")
Central_South_Asia <- subset(refs_numbered_ID_pop_Q, pop == "Central South Asia (HGDP)")
South_Asia_SGDP2 <- subset(refs_numbered_ID_pop_Q, pop == "South Asia (SGDP),Central South Asia (HGDP)")
South_Asia_SGDP3 <- subset(refs_numbered_ID_pop_Q, pop == "South Asia (SGDP)")
South_Asia_SGDP_Central_South_Asia_HGDP1 <- subset(refs_numbered_ID_pop_Q, pop == "Central South Asia (HGDP)")
South_Asia_SGDP_Central_South_Asia_HGDP2 <- subset(refs_numbered_ID_pop_Q, pop == "South Asia (SGDP),Central South Asia (HGDP)")
SAS <- rbind.fill(South_Asian, South_Asia_SGDP, Central_South_Asia, South_Asia_SGDP2, South_Asia_SGDP3, South_Asia_SGDP_Central_South_Asia_HGDP1,South_Asia_SGDP_Central_South_Asia_HGDP2)

African <- subset(refs_numbered_ID_pop_Q, pop == "African Ancestry")
African_SGDP <- subset(refs_numbered_ID_pop_Q, pop == "African Ancestry,Africa (SGDP)")
Africa_HGDP <- subset(refs_numbered_ID_pop_Q, pop == "Africa (HGDP)")
African_SGDP2 <- subset(refs_numbered_ID_pop_Q, pop == "Africa (SGDP)")
African_SGDP_HGDP <- subset(refs_numbered_ID_pop_Q, pop == "Africa (SGDP),Africa (HGDP)")
AFR <- rbind.fill(Africa_HGDP, African, African_SGDP, African_SGDP_HGDP, African_SGDP2)

Oceania_HGDP_SGDP <- subset(refs_numbered_ID_pop_Q, pop == "Oceania (SGDP),Oceania (HGDP)")
Oceania_HGDP <- subset(refs_numbered_ID_pop_Q, pop == "Oceania (HGDP)")
Oceania_SGDP <- subset(refs_numbered_ID_pop_Q, pop == "Oceania (SGDP)")
OCE <- rbind.fill(Oceania_HGDP, Oceania_HGDP_SGDP, Oceania_SGDP)

Middle_East_HGDP <- subset(refs_numbered_ID_pop_Q, pop == "Middle East (HGDP)")
Middle_East_HGDP_Africa_SGDP <- subset(refs_numbered_ID_pop_Q, pop == "Middle East (HGDP),Africa (SGDP)")
West_Eurasia <- subset(refs_numbered_ID_pop_Q, pop == "West Eurasia (SGDP)")
MDE <- rbind.fill(Middle_East_HGDP, Middle_East_HGDP_Africa_SGDP, West_Eurasia)

Central_Asia_and_Siberia_SGDP <- subset(refs_numbered_ID_pop_Q, pop == "Central Asia and Siberia (SGDP)")

EUR <- EUR[,c(2:8)]
EUR[EUR$pop == "European Ancestry",]$pop = "EUR"
EUR[EUR$pop == "Europe (HGDP)",]$pop = "EUR"
EUR[EUR$pop == "European Ancestry,West Eurasia (SGDP)",]$pop = "EUR"

AFR <- AFR[,c(2:8)]
AFR[AFR$pop == "Africa (HGDP)",]$pop = "AFR"
AFR[AFR$pop == "African Ancestry",]$pop = "AFR"
AFR[AFR$pop == "African Ancestry,Africa (SGDP)",]$pop = "AFR"
AFR[AFR$pop == "Africa (SGDP)",]$pop = "AFR"
AFR[AFR$pop == "Africa (SGDP),Africa (HGDP)",]$pop = "AFR"

AMR <- AMR[,c(2:8)]
AMR[AMR$pop == "America (HGDP)",]$pop = "AMR"
AMR[AMR$pop == "American Ancestry",]$pop = "AMR"
AMR[AMR$pop == "America (SGDP),America (HGDP)",]$pop = "AMR"
AMR[AMR$pop == "America (SGDP)",]$pop = "AMR"

SAS <- SAS[,c(2:8)]
SAS[SAS$pop == "South Asian Ancestry",]$pop = "SAS"
SAS[SAS$pop == "Central South Asia (HGDP)",]$pop = "SAS"
SAS[SAS$pop == "South Asia (SGDP),South Asian Ancestry",]$pop = "SAS"
SAS[SAS$pop == "South Asia (SGDP)",]$pop = "SAS"

EAS <- EAS[,c(2:8)]
EAS[EAS$pop == "East Asian Ancestry",]$pop = "EAS"
EAS[EAS$pop == "East Asia (SGDP),East Asian Ancestry",]$pop = "EAS"
EAS[EAS$pop == "East Asia (HGDP)",]$pop = "EAS"
EAS[EAS$pop == "East Asia (SGDP)",]$pop = "EAS"

OCE <- OCE[,c(2:8)]
OCE[OCE$pop == "Oceania (HGDP)",]$pop = "OCE"
OCE[OCE$pop == "Oceania (SGDP),Oceania (HGDP)",]$pop = "OCE"
OCE[OCE$pop == "Oceania (SGDP)",]$pop = "OCE"

MDE <- MDE[,c(2:8)]
MDE[MDE$pop == "Middle East (HGDP)",]$pop = "MDE"
MDE[MDE$pop == "West Eurasia (SGDP)",]$pop = "MDE"
MDE[MDE$pop == "Middle East (HGDP),Africa (SGDP)",]$pop = "MDE"

CAS <- Central_Asia_and_Siberia_SGDP[,c(2:8)]
CAS[CAS$pop == "Central Asia and Siberia (SGDP)",]$pop = "CAS"

temp2 <- rbind(EUR,AFR,AMR,SAS,EAS,MDE,OCE,CAS)
raw_fam2 <- unique(raw_fam[,c(1,3)])
temp3 <- merge (temp2,raw_fam2,by="WES_ID")
temp3 <- unique(temp3[,c(8,1,2)])
temp4 <- merge(temp3,raw_fam2,by="WES_ID", all = TRUE)
temp4 <- temp4[,c(4,1,3)]
write.csv(temp4,"admixture_supervised.pop")
```
```{r}
White <- unique(subset(samples_numbered_ID_race_Q, pop == "White"))
```
```{r}
Asian <- unique(subset(samples_numbered_ID_race_Q, pop == "Asian"))     
```
```{r}
Black <- unique(subset(samples_numbered_ID_race_Q, pop == "Black_or_African_American"))            
```
```{r}
Other <- unique(subset(samples_numbered_ID_race_Q, pop == "Other"))    
```
```{r}
PacIsl <- unique(subset(samples_numbered_ID_race_Q, pop == "Hawaiian_or_Other_Pacific_Islander"))
```
```{r}
AIAN <- unique(subset(samples_numbered_ID_race_Q, pop == "American_Indian_or_Alaska_Native"))
```
```{r}
UK1 <- unique(subset(samples_numbered_ID_race_Q, pop == "Unknown/Not_Reported"))
```
```{r}
UK2 <- unique(subset(samples_numbered_ID_race_Q, pop == "Unknown"))
```
```{r}
UKrace <- unique(rbind(UK1,UK2))       
```
```{r}
NHL <- unique(subset(samples_numbered_ID_ethnicity_Q, pop == "Not_Hispanic/Not_Latino"))
HL <- unique(subset(samples_numbered_ID_ethnicity_Q, pop == "Hispanic/Latino"))            
UK3 <- unique(subset(samples_numbered_ID_ethnicity_Q, pop == "Unknown/Not_Reported"))     
UK4 <- unique(subset(samples_numbered_ID_ethnicity_Q, pop == "Unknown")) 
UKeth <- unique(rbind(UK3,UK4))
```
```{r}
write.csv(NHL, "IDs_ORIEN_TCGA_NHL_pop_Q_K5.csv")
write.csv(HL, "IDs_ORIEN_TCGA_HL_Q_K5.csv")
write.csv(UKeth, "IDs_ORIEN_UnknownEthnicity_race_Q_K5.csv")
write.csv(White, "IDs_ORIEN_TCGA_White_Q_K5.csv")
write.csv(Asian, "IDs_ORIEN_TCGA_Asian_Q_K5.csv")
write.csv(Black, "IDs_ORIEN_TCGA_Black_Q_K5.csv")
write.csv(Other, "IDs_ORIEN_TCGA_OtherRace_Q_K5.csv")
write.csv(PacIsl, "IDs_ORIEN_TCGA_PacificIslander_Q_K5.csv")
write.csv(AIAN, "IDs_ORIEN_TCGA_AmericanIndian_Q_K5.csv")
write.csv(UKrace, "IDs_ORIEN_TCGA_UnknownRace_Q_K5.csv")

write.csv(EUR, "IDs_HGDP_1000GP_EUR_Q_K5.csv")
write.csv(AFR, "IDs_HGDP_1000GP_AFR_Q_K5.csv")
write.csv(AMR, "IDs_HGDP_1000GP_AMR_Q_K5.csv")
write.csv(SAS, "IDs_HGDP_1000GP_SAS_Q_K5.csv")
write.csv(EAS, "IDs_HGDP_1000GP_EAS_Q_K5.csv")
write.csv(OCE, "IDs_HGDP_1000GP_OCE_Q_K5.csv")
write.csv(MDE, "IDs_HGDP_1000GP_MDE_Q_K5.csv")

NHL2 <- NHL[,c(2:8)]; HL2 <- HL[,c(2:8)]; UKeth2 <- UKeth[,c(2:8)]
refs_pop_samples_ethnicity_Q <- rbind(NHL2, HL2, UKeth2, EUR, AFR, AMR, SAS, EAS, OCE, MDE)
write.csv(refs_pop_samples_ethnicity_Q, "HGDP_1000GP_pops_ORIEN_TCGA_ethnicities_plusQ_K5.csv")
White2 <- White[,c(2:8)]; Asian2 <- Asian[,c(2:8)]; Black2 <- Black[,c(2:8)]; Other2 <- Other[,c(2:8)]; PacIsl2 <- PacIsl[,c(2:8)]; AIAN2 <- AIAN[,c(2:8)]; UKrace2 <- UKrace[,c(2:8)]
refs_pop_samples_race_Q <- rbind(White2, Asian2, Black2, Other2, PacIsl2, AIAN2, UKrace2, EUR, AFR, AMR, SAS, EAS, OCE, MDE)
write.csv(refs_pop_samples_ethnicity_Q, "HGDP_1000GP_pops_ORIEN_TCGA_races_plusQ_K5.csv")

```

1) Reference populations:

AFR
```{r}
df <- AFR
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "population"
names(df)[3] <- "Q1"
names(df)[4] <- "Q2"
names(df)[5] <- "Q3"
names(df)[6] <- "Q4"
names(df)[7] <- "Q5"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}

plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ population, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ population, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_1000GP_and_HGDP_AFR_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

AMR
```{r}
df <- AMR
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "population"
names(df)[3] <- "Q1"
names(df)[4] <- "Q2"
names(df)[5] <- "Q3"
names(df)[6] <- "Q4"
names(df)[7] <- "Q5"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ population, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ population, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_1000GP_and_HGDP_AMR_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

#CAS
```{r}
df <- CAS
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "population"
names(df)[3] <- "Q1"
names(df)[4] <- "Q2"
names(df)[5] <- "Q3"
names(df)[6] <- "Q4"
names(df)[7] <- "Q5"

# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ population, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ population, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_1000GP_and_HGDP_CAS_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

EAS
```{r}
df <- EAS
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "population"
names(df)[3] <- "Q1"
names(df)[4] <- "Q2"
names(df)[5] <- "Q3"
names(df)[6] <- "Q4"
names(df)[7] <- "Q5"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ population, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ population, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_1000GP_and_HGDP_EAS_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

EUR
```{r}
df <- EUR
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "population"
names(df)[3] <- "Q1"
names(df)[4] <- "Q2"
names(df)[5] <- "Q3"
names(df)[6] <- "Q4"
names(df)[7] <- "Q5"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ population, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ population, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_1000GP_and_HGDP_EUR_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

MDE
```{r}
df <- MDE
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "population"
names(df)[3] <- "Q1"
names(df)[4] <- "Q2"
names(df)[5] <- "Q3"
names(df)[6] <- "Q4"
names(df)[7] <- "Q5"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ population, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ population, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_1000GP_and_HGDP_MDE_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

OCE
```{r}
df <- OCE
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "population"
names(df)[3] <- "Q1"
names(df)[4] <- "Q2"
names(df)[5] <- "Q3"
names(df)[6] <- "Q4"
names(df)[7] <- "Q5"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ population, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ population, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_1000GP_and_HGDP_OCE_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

SAS
```{r}
df <- SAS
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "population"
names(df)[3] <- "Q1"
names(df)[4] <- "Q2"
names(df)[5] <- "Q3"
names(df)[6] <- "Q4"
names(df)[7] <- "Q5"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ population, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ population, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_1000GP_and_HGDP_SAS_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

#WEU
```{r}
df <- WEU
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "population"
names(df)[3] <- "Q1"
names(df)[4] <- "Q2"
names(df)[5] <- "Q3"
names(df)[6] <- "Q4"
names(df)[7] <- "Q5"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")

plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ population, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ population, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_1000GP_and_HGDP_WEU_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```


2) ORIEN +TCGA by self-reported race

White
```{r}
df <- White
df <- unique(df)
df <- df[,c(2:8)]

#df$sum=rowSums(cbind(df$`Q1 - EAS`,df$`Q2 - EUR`,df$`Q3 - AMR`,df$`Q4 - SAS`,df$`Q5 - AFR`),na.rm=TRUE)

# Add id and self_ID_race label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_race"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"

# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_race"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_race, drop=TRUE, space="free") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))
#   facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_y") +
#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=mdat$Ancestry) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_race_White_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

Black or African American
```{r}
df <- Black
df <- unique(df)
df <- df[,c(2:8)]

# Add id and self_ID_race label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_race"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"

# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_race"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_race_Black_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

Asian
```{r}
df <- Asian
df <- unique(df)
df <- df[,c(2:8)]

# Add id and self_ID_race label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_race"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_race"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_race_Asian_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

American Indian or Alaska Native
```{r}
df <- AIAN
df <- unique(df)
df <- df[,c(2:8)]

# Add id and self_ID_race label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_race"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_race"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_race_AIAN_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

Hawaiian or Other Pacific Islander
```{r}
df <- PacIsl
df <- unique(df)
df <- df[,c(2:8)]

# Add id and self_ID_race label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_race"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"



# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_race"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_race_PacIsl_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

Unknown
```{r}
df <- UKrace
df <- unique(df)
df <- df[,c(2:8)]

# Add id and self_ID_race label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_race"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_race"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_race_Unknown_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

Other
```{r}
df <- Other
df <- unique(df)
df <- df[,c(2:8)]

# Add id and self_ID_race label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_race"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_race"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_race, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_race_Other_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

check NAs!


3) ORIEN +TCGA by self-reported Ethnicity

Hispanic/Latino
```{r}
df <- HL
df <- unique(df)
df <- df[,c(2:8)]

# Add id and self_ID_ethnicity label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_ethnicity"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_ethnicity"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_ethnicity, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ self_ID_ethnicity, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_ethnicity, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_ethnicity_HispanicLatino_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

Not Hispanic/Latino
```{r}
df <- NHL
df <- unique(df)
df <- df[,c(2:8)]

# Add id and self_ID_ethnicity label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_ethnicity"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_ethnicity"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_ethnicity, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ self_ID_ethnicity, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_ethnicity, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_ethnicity_NOTHispanicLatino_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

Unknown
```{r}
df <- UKethnicity
df <- unique(df)
df <- df[,c(2:8)]

# Add id and self_ID_ethnicity label columns. Needed for reshape2::melting and plotting.
names(df)[1] <- "ID"
names(df)[2] <- "self_ID_ethnicity"
names(df)[3] <- "Q1 - EAS"
names(df)[4] <- "Q2 - EUR"
names(df)[5] <- "Q3 - AMR"
names(df)[6] <- "Q4 - SAS"
names(df)[7] <- "Q5 - AFR"


# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df, id.vars=c("ID", "self_ID_ethnicity"), 
            variable.name="Ancestry", value.name="Fraction")
```
```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ self_ID_ethnicity, drop=TRUE, space="free", scales="free_x") +
#   facet_grid(. ~ self_ID_ethnicity, drop=TRUE, space="free", scales="free_y") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0))

#col8 = c("crimson", "gold", "yellow", "forestgreen", "dodgerblue", "navyblue", "darkviolet", "hotpink")
#names(col8) = levels(mdat$Ancestry)
p
# Customized stacked bar plot:
# Sort ancestry order by overall 'abundance' of each ancestry.
#mdat$Ancestry = factor(mdat$Ancestry, 
#                       levels=names(sort(colSums(df[, 1:17]), decreasing=TRUE)))
# Colors taken from:
# https://github.com/mbostock/d3/wiki/Ordinal-Scales#category20
#col8 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
#          "#8c564b", "#e377c2", "#7f7f7f")
#p2 = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
#     geom_bar(stat="identity", position="fill", width=.5, colour="grey25") +
#     facet_grid(. ~ self_ID_ethnicity, drop=TRUE, space="free", scales="free") +
#     theme(panel.grid=element_blank()) +
#     theme(panel.background=element_rect(fill=NA, colour="grey25")) +
#     theme(panel.spacing.x=grid:::unit(0.5, "lines")) +
#     theme(axis.title.x=element_blank()) +
#     theme(axis.text.x=element_blank()) +
#     theme(axis.ticks.x=element_blank()) +
#     theme(strip.background=element_blank()) +
#     theme(strip.text=element_text(size=12)) +
#     theme(legend.key=element_rect(colour="grey25")) +
#     scale_x_discrete(expand=c(0, 0)) +
#     scale_y_continuous(expand=c(0, 0)) +
#     scale_fill_manual(values=col8) +
#     guides(fill=guide_legend(override.aes=list(colour=NULL)))

png("bar_plots_ORIEN+TCGA_by_ethnicity_Unknown_K5.png", width=10, height=5, units="in", res=1000)
#grid.arrange(p, nrow=1)
p
dev.off()

```

check NAs!

