---
title: "Admixture_Stacked_Barplot_K5"
output: html_notebook
---

Load libraries
```{r}
library(ggplot2)
library(reshape2)
library(gridExtra)
library(dplyr)
library(data.table)
library(tidyverse)
library(plotly)
library(plyr)
library(stringi)
```

Read in race/ethnicity info from ORIEN 
```{r}
clinlinkage <- read.csv("20PRJ071USC_20210928_ClinicalMolLinkage_V4.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
key_SLID <- subset(clinlinkage, Tumor.Germline == "Germline")
key_SLID2 <- key_SLID[,c(1,3)]
demos <- read.csv("20PRJ071USC_20210928_PatientMaster_V4.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
demos2 <- demos[,c(1,5,6)]
names(key_SLID2)[names(key_SLID2 ) == "ORIENAvatarKey" ] <- "AvatarKey"
merge <- merge(x = key_SLID2,y = demos2, by='AvatarKey', sort=FALSE, all=TRUE)
names(merge)[names(merge ) == "WES" ] <- "WES_ID"
merge2 <- merge[,c(2,3,4)]

USC <-  read.csv("20211203_USC_ClinicalMolLinkage_V4.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
names(USC)[names(USC ) == "ORIENAvatarKey" ] <- "key"
key <- read.csv("V4_RecordKey_Jan2022.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
names(key)[names(key ) == "Avatar.Patient.ID" ] <- "key"
key <- key[,c(4,6,7)]
USC2 <- merge(x=USC, y=key, by="key")
USC2 <- USC2[,c(4,21,22)]
names(USC2)[names(USC2 ) == "WES" ] <- "WES_ID"
all <- bind_rows(merge2,USC2)
all <- all[-which(all$WES_ID == ""), ]

all[all$Race == "Black",]$Race = "Black_or_African_American"
all[all$Race == "Unknown/Not Reported",]$Race = "Unknown/Not_Reported"
all[all$Race == "Other Asian, including Asian, NOS and Oriental, NOS",]$Race = "Asian"
all[all$Race == "Filipino",]$Race = "Asian"
all[all$Race == "Chinese",]$Race = "Asian"
all[all$Race == "Korean",]$Race = "Asian"
all[all$Race == "American Indian, Aleutian, or Eskimo (includes all indigenous populations of the Western Hemisphere)",]$Race = "American_Indian_or_Alaska_Native"
all[all$Race == "American Indian, Aleutian, or Eskimo (includes all indigenous populations of the Western hemisphere)",]$Race = "American_Indian_or_Alaska_Native"
all[all$Race == "Asian Indian or Pakistani, NOS",]$Race = "Asian"
all[all$Race == "Asian Indian",]$Race = "Asian"
all[all$Race == "Vietnamese",]$Race = "Asian"
all[all$Race == "Japanese",]$Race = "Asian"
all[all$Race == "Pacific Islander, NOS",]$Race = "Hawaiian_or_Other_Pacific_Islander"
all[all$Race == "Unknown/Not Applicable",]$Race = "Unknown/Not_Reported"
all[all$Race == "Hawaiian",]$Race = "Hawaiian_or_Other_Pacific_Islander"
all[all$Race == "Laotian",]$Race = "Asian"
all[all$Race == "Pakistani",]$Race = "Asian"
all[all$Race == "Kampuchean (Cambodian)",]$Race = "Asian"
all[all$Race == "Polynesian, NOS",]$Race = "Hawaiian_or_Other_Pacific_Islander"
all[all$Race == "Thai",]$Race = "Asian"
all[all$Race == "Guamanian, NOS",]$Race = "Hawaiian_or_Other_Pacific_Islander"

all[all$Ethnicity == "Non-Spanish; Non-Hispanic",]$Ethnicity = "Not_Hispanic/Not_Latino" 
all[all$Ethnicity == "Spanish, NOS; Hispanic, NOS; Latino, NOS",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Unknown/Not Reported",]$Ethnicity = "Unknown/Not_Reported"
all[all$Ethnicity == "Mexican (includes Chicano)",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "South or Central American (except Brazil)",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Puerto Rican",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Unknown/Not Applicable",]$Ethnicity = "Unknown/Not_Reported"
all[all$Ethnicity == "Cuban",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Other specified Spanish/Hispanic origin (includes European; excludes Dominican Republic)",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Spanish surname only",]$Ethnicity = "Hispanic/Latino"
all[all$Ethnicity == "Dominican Republic",]$Ethnicity = "Hispanic/Latino"


TCGA_race <- read.csv("TCGA_vcfIDs_race_uniq_final.tsv", sep = "\t", header = FALSE, stringsAsFactors = FALSE, fill = TRUE)
TCGA_ethnicity <- read.csv("TCGA_vcfIDs_ethnicity_uniq_final.tsv", sep = "\t", header = FALSE, stringsAsFactors = FALSE, fill = TRUE)
names(TCGA_race)[2] <- "Race"
names(TCGA_ethnicity)[2] <- "Ethnicity"
names(TCGA_race)[1] <- "WES_ID"
names(TCGA_ethnicity)[1] <- "WES_ID"
TCGA <- merge(x = TCGA_race, y = TCGA_ethnicity, by="WES_ID")

merge5 <- bind_rows(TCGA,all)

race <- merge5[,c(1,2)]
names(race)[names(race) == "Race"] <- "pop"
ethnicity <- merge5[,c(1,3)]
names(ethnicity)[names(ethnicity) == "Ethnicity"] <- "pop"

str_replace(race$WES_ID, "_st_g", "")
str_replace(ethnicity$WES_ID, "_st_g", "")
```

Standardize/harmonize and combine fam + Q + pop info
```{r}
#Q1 <- read.table("gnomad.7.Q", head = FALSE,  sep = " ", stringsAsFactors = FALSE)

Q1 <- read.table("refs.7.Q", head = FALSE,  sep = " ", stringsAsFactors = FALSE)
raw_fam1 <- read.table("gnomad.fam", head = FALSE, sep = " ", stringsAsFactors = FALSE)

gnomAD <- cbind(Q1,raw_fam1)
gnomAD <- gnomAD[,c(8,1:7)]
names(gnomAD)[1] <- "WES_ID"
names(gnomAD)[2] <- "Q1"
names(gnomAD)[3] <- "Q2"
names(gnomAD)[4] <- "Q3"
names(gnomAD)[5] <- "Q4"
names(gnomAD)[6] <- "Q5"
names(gnomAD)[7] <- "Q6"
names(gnomAD)[8] <- "Q7"

```

```{r}
#refpops_HGDP <- read_tsv("igsr-human genome diversity project.tsv.tsv")
#refpops_1000GP <- read_tsv("igsr-1000 genomes 30x on grch38.tsv.tsv")
#refpops_igsr <- read_tsv("igsr_samples3.tsv")

#refpops_igsr <- refpops_igsr[,c(1,4)]
#refpops_HGDP2 <- refpops_HGDP[,c(1,4)]
#refpops_1000GP2 <- refpops_1000GP[,c(1,4)]
#refpops2 <- unique(rbind(refpops_HGDP2,refpops_1000GP2,refpops_igsr))
refpops2 <- read.csv("ORIEN_gnomad_IDs_pops.csv")
#names(refpops2)[names(refpops2) == "Sample name"] <- "WES_ID"
#names(refpops2)[names(refpops2) == "Superpopulation name"] <- "pop"
#GP1000_PR <- subset(refpops2,refpops2$`Population code`=="PUR")

#GP1000_PR2 <- merge(GP1000_PR, HGDP_1000GP, by="WES_ID", all = FALSE)
df <- merge(refpops2,gnomAD, by="WES_ID")
gnomAD_PR <- subset(df,df$pop =="Puerto Rican")
```

Key
#000000 = black --> EUR N.
#E69F00 = light orange --> Indig Am.
#56B4E9 = light blue --> OCE
#009E73 = green --> EUR South
#F0E442 = yellow --> SAS
#0072B2 = dark blue --> AFR C/S
#D55E00 = red/dark orange --> EAS
#CC79A7 = pink --> AFR W.

In this run:
Q1 = EUR N
Q2 = AMR
Q3 = OCE
Q4 = MDE/EUR S
Q5 = SAS
Q6 = AFR C/S
Q7 = EAS
Q8 = AFR W


```{r}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

gnomAD PR
```{r}
#df <- GP1000_PR2
df1 <- unique(gnomAD_PR)

df2 <- df1[,c(1,3,5:11)]
names(df2)[1] <- "ID"
names(df2)[2] <- "gnomAD_PR"
names(df2)[3] <- "Q1"
names(df2)[4] <- "Q2"
names(df2)[5] <- "Q3"
names(df2)[6] <- "Q4"
names(df2)[7] <- "Q5"
names(df2)[8] <- "Q6"
names(df2)[9] <- "Q7"


mdat = reshape2::melt(df2, id.vars=c("ID", "gnomAD_PR"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]
```

```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
        geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ gnomAD_PR, drop=TRUE, space="free") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("bar_plots_gnomAD_PR_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```


ORIEN: Puerto Rican
```{r}
#Q2 <- read.table("ORIEN.7.Q", head = FALSE,  sep = " ", stringsAsFactors = FALSE)
Q2 <- read.table("samples.7.Q", head = FALSE,  sep = " ", stringsAsFactors = FALSE)
raw_fam2 <- read.table("ORIEN.fam", head = FALSE, sep = " ", stringsAsFactors = FALSE)
raw_fam2 <- raw_fam2[,c(1,2)]
names(raw_fam2)[1] <- "WES_ID"

samples <- cbind(Q2,raw_fam2)
samples <-samples[,c(8,1:7)]
names(samples)[1] <- "WES_ID"
names(samples)[2] <- "Q1"
names(samples)[3] <- "Q2"
names(samples)[4] <- "Q3"
names(samples)[5] <- "Q4"
names(samples)[6] <- "Q5"
names(samples)[7] <- "Q6"
names(samples)[8] <- "Q7"

df3 <- read.csv("ORIEN_PR_list.csv")
ORIEN_PR <- merge(df3, samples, by="WES_ID")
```

```{r}
df4 <- unique(ORIEN_PR)

#df4$sum=rowSums(cbind(df4$`Q1 - EAS`,df4$`Q2 - EUR`,df4$`Q3 - AMR`,df4$`Q4 - SAS`,df4$`Q5 - AFR`),na.rm=TRUE)

# Add id and self_ID_race label columns. Needed for reshape2::melting and plotting.
names(df4)[1] <- "ID"
names(df4)[2] <- "ORIEN_PR"
names(df4)[3] <- "Q1"
names(df4)[4] <- "Q2"
names(df4)[5] <- "Q3"
names(df4)[6] <- "Q4"
names(df4)[7] <- "Q5"
names(df4)[8] <- "Q6"
names(df4)[9] <- "Q7"

# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(df4, id.vars=c("ID", "ORIEN_PR"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]
```


```{r}
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


```{r}
plot.new()            # Create empty plot in RStudios default window
dev.new(width = 400,   # Create new plot window
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

# Simple stacked bar plot:
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry, order=Ancestry)) +
        geom_bar(stat="identity", position="stack") +
    facet_grid(. ~ ORIEN_PR, drop=TRUE, space="free") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=12)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("bar_plots_ORIEN_PR_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```

#--------------
gnomAD

```{r}
#earlier step: df <- merge(refpops2,gnomAD, by="WES_ID")
AFR <- subset(df, superpop == "AFR")
#order below is C,C,E,E,S,S,S,S,S,W,W,W,W,W,x,x AFR
AFR$pop <- factor(AFR$pop, levels = c("Mbuti","Biaka","Luhya","Bantu Kenya", "San","Bantu Tswana", "Bantu Herero", "Ju_hoan_North", "Mende","Mandenka", "Gambian Mandinka", "Yoruba","Esan","African Caribbean", "African Ancestry SW"), ordered = TRUE)
AFR <- AFR[order(AFR$pop),]
write.csv(AFR,"gnomAD_AFR_ordered_K5.csv")
AFR <- AFR[,c(1,3,5:11)]

AMR <- subset(df, superpop == "AMR")
write.csv(AMR,"gnomAD_AMR_K5.csv")
AMR <- AMR[,c(1,3,5:11)]

EAS <- subset(df, superpop == "EAS")
write.csv(EAS,"gnomAD_EAS_K5.csv")
EAS <- EAS[,c(1,3,5:11)]

SAS <- subset(df, superpop == "SAS")
write.csv(SAS,"gnomAD_SAS_K5.csv")
SAS <- SAS[,c(1,3,5:11)]

OCE <- subset(df, superpop == "OCE")
OCE <- OCE[,c(1,3,5:11)]

MDE <- subset(df, superpop == "MDE")
MDE <- MDE[,c(1,3,5:11)]

EUR <- subset(df, superpop == "EUR")
#order below so that Northern EUR --> Southern EUR, and end w/ EUR in US
EUR$pop <- factor(EUR$pop, levels = c("British","Finnish","Orcadian","Russian","Adygei","French","Basque","Iberian","Bergamo Italian","Tuscan","Toscani","Sardinian","BedouinB","CEPH"), ordered = TRUE)
EUR <- EUR[order(EUR$pop),]
write.csv(EUR,"gnomAD_EUR_ordered_K5.csv")
EUR <- EUR[,c(1,3,5:11)]

CAS <- subset(df, superpop == "CAS")
CAS <- CAS[,c(1,3,5:11)]
```

AFR
```{r}
names(AFR)[1] <- "ID"
names(AFR)[2] <- "population"
names(AFR)[3] <- "Q1"
names(AFR)[4] <- "Q2"
names(AFR)[5] <- "Q3"
names(AFR)[6] <- "Q4"
names(AFR)[7] <- "Q5"
names(AFR)[8] <- "Q6"
names(AFR)[9] <- "Q7"

mdat = reshape2::melt(AFR, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]

```

```{r}
plot.new()            
dev.new(width = 400,  
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack", colour="black", size=0.005) +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=6)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("barplots_gnomAD_AFR_admixture_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```

AMR
```{r}
AMR <- unique(AMR)
names(AMR)[1] <- "ID"
names(AMR)[2] <- "population"
names(AMR)[3] <- "Q1"
names(AMR)[4] <- "Q2"
names(AMR)[5] <- "Q3"
names(AMR)[6] <- "Q4"
names(AMR)[7] <- "Q5"
names(AMR)[8] <- "Q6"
names(AMR)[9] <- "Q7"

mdat = reshape2::melt(AMR, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]
```

```{r}
plot.new()            
dev.new(width = 400,  
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack", colour="black", size=0.005) +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=6)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("barplots_gnomAD_AMR_admixture_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```

CAS
```{r}
CAS <- unique(CAS)
# Add id and population label columns. Needed for reshape2::melting and plotting.
names(CAS)[1] <- "ID"
names(CAS)[2] <- "population"
names(CAS)[3] <- "Q1"
names(CAS)[4] <- "Q2"
names(CAS)[5] <- "Q3"
names(CAS)[6] <- "Q4"
names(CAS)[7] <- "Q5"
names(CAS)[8] <- "Q6"
names(CAS)[9] <- "Q7"

# Melt (reshape data from wide format to long format).
mdat = reshape2::melt(CAS, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]
```

```{r}
plot.new()           
dev.new(width = 400,   
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)
p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack", colour="black", size=0.005) +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=6)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("barplots_gnomAD_CAS_admixture_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```

EAS
```{r}
EAS <- unique(EAS)
names(EAS)[1] <- "ID"
names(EAS)[2] <- "population"
names(EAS)[3] <- "Q1"
names(EAS)[4] <- "Q2"
names(EAS)[5] <- "Q3"
names(EAS)[6] <- "Q4"
names(EAS)[7] <- "Q5"
names(EAS)[8] <- "Q6"
names(EAS)[9] <- "Q7"

mdat = reshape2::melt(EAS, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]
```

```{r}
plot.new()        
dev.new(width = 400,  
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack", colour="black", size=0.005) +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=6)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("barplots_gnomAD_EAS_admixture_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```

EUR
```{r}
EUR <- unique(EUR)
names(EUR)[1] <- "ID"
names(EUR)[2] <- "population"
names(EUR)[3] <- "Q1"
names(EUR)[4] <- "Q2"
names(EUR)[5] <- "Q3"
names(EUR)[6] <- "Q4"
names(EUR)[7] <- "Q5"
names(EUR)[8] <- "Q6"
names(EUR)[9] <- "Q7"

mdat = reshape2::melt(EUR, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]
```

```{r}
plot.new()          
dev.new(width = 400,  
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack", colour="black", size=0.005) +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=6)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("barplots_gnomAD_EUR_admixture_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```

MDE
```{r}
MDE <- unique(MDE)
names(MDE)[1] <- "ID"
names(MDE)[2] <- "population"
names(MDE)[3] <- "Q1"
names(MDE)[4] <- "Q2"
names(MDE)[5] <- "Q3"
names(MDE)[6] <- "Q4"
names(MDE)[7] <- "Q5"
names(MDE)[8] <- "Q6"
names(MDE)[9] <- "Q7"


mdat = reshape2::melt(MDE, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]
```


```{r}
plot.new()       
dev.new(width = 400,  
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack", colour="black", size=0.005) +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=6)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("barplots_gnomAD_MDE_admixture_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```

OCE
```{r}
OCE <- unique(OCE)
names(OCE)[1] <- "ID"
names(OCE)[2] <- "population"
names(OCE)[3] <- "Q1"
names(OCE)[4] <- "Q2"
names(OCE)[5] <- "Q3"
names(OCE)[6] <- "Q4"
names(OCE)[7] <- "Q5"
names(OCE)[8] <- "Q6"
names(OCE)[9] <- "Q7"


mdat = reshape2::melt(OCE, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]
```

```{r}
plot.new()          
dev.new(width = 400,  
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack", colour="black", size=0.005) +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=6)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("barplots_gnomAD_OCE_admixture_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```

SAS
```{r}
SAS <- unique(SAS)
names(SAS)[1] <- "ID"
names(SAS)[2] <- "population"
names(SAS)[3] <- "Q1"
names(SAS)[4] <- "Q2"
names(SAS)[5] <- "Q3"
names(SAS)[6] <- "Q4"
names(SAS)[7] <- "Q5"
names(SAS)[8] <- "Q6"
names(SAS)[9] <- "Q7"

mdat = reshape2::melt(SAS, id.vars=c("ID", "population"), 
            variable.name="Ancestry", value.name="Fraction")

mdat = mdat[
  with(mdat, order(mdat$Fraction, mdat$Ancestry)),
]
```

```{r}
plot.new()            
dev.new(width = 400,  
        height = 600,
        unit = "px",
        noRStudioGD = TRUE)

p = ggplot(mdat, aes(x=ID, y=Fraction, fill=Ancestry)) +
    geom_bar(stat="identity", position="stack", colour="black", size=0.005) +
    facet_grid(. ~ population, drop=TRUE, space="free", scales="free_x") +
    theme(panel.spacing.x=grid:::unit(0.01, "lines")) +
    theme(strip.text=element_text(size=6)) +
    scale_x_discrete(expand=c(0, 0)) + scale_fill_manual(values=cbPalette)

p

png("barplots_gnomAD_SAS_admixture_K7.png", width=10, height=5, units="in", res=1000)
p
dev.off()

```

#WEU - combined w/ Middle East (MDE)
#--------------

